|

### 2022-01-18 text/ManuscriptDraft.Rmd
Rev1 comments from TTH and AI implmented or comment added

A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track. Often simple statistics are used to describe the flowering phenology of a population of plants, e.g. onset or peak of flowering. Similarly, flower visitation rates are based on crude observation of flower populations. Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals. Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series. We compare the results of automatic tracking with manual tracking. Individual flower tracking could facilitate finegrained studies on plant reproductive success.
---

@@

Here, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.

We present and evaluate an automatic flower tracking and filtering algorithm. Many tracking algorithms, such as the Simple online and realtime tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. We therefore base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.

For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach. The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors. The approach allows for confident upscaling of the method. I.e., when running the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.

@@

`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software and assigned each individual flower a unique ID. These annotations constitute our ground truth tracks.

@@

The tracking algorithm tracks objects based on distances between centroids of bounding boxes. The algorithm has a set of user adjusted parameters that can optimize tracking accuracy. The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers. It is important to note, however, that the tracking algorithm could be used to track any objects. The algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame). The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked. The method is fast, however. Tracking of a series containing 85 objects ran at ~0.02 seconds per frame on a standard laptop.

Several things complicate the task of tracking individual flowers through the time-lapse series. First, as the wind shifts, the flower heads changes direction. This can happen instantaneously (i.e. between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other. With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.

Second, when the location of the flowers change they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view. This can cause a track to be lost and a new track erroneously being initiated when the flower reappears. However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area. Concurrently, this deals with potential false negatives. If a given flower has not been annotated in a few frames, the track will not be lost. The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.

@@

#
![](../figures/figure_1.png){ width=100% }

@@

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we ran the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3.179 combinations): Values for max disappeared and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100. Note that set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches snd compare the tracking results between optimal settings and all parameters set to zero.

`r word_comment("I haven't included these results. Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground tracks. First, we calculated the largest distance between any two points within any track for any flower in each of the five series. Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.

#
## Evaluating tracking perfomance

The optimal way of quantifying tracking performance depends on the goal of the tracking. To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct. To derive flowering length, in theory we just need to track the most extreme points correctly and can ignore the intermediate points. Lastly, we may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.

The multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We calculate the number of mismatches and the MOTA score to evaluate performance of our tracking and filtering algorithms. Further, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal.

@@

We base our filtering on a density based clustering of the centroid of a geometry of each track. The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid. For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line. For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle. For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon. For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.

We then apply the filtering based on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches. The filtering is done in two steps. First, the DBSCAN clustering algorithm is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together. Second, all tracks that were not assigned to a unique cluster are removed. Tracks that are spatially isolated remains. We evaluate the tracking accuracy of the remaining tracks. We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.

@@

The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig. 4. Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max disappeared and max distance, respectively).

@@

We ran our tracking algorithm on 3xxx combination of the three parameters. Using the parameters increased the performance of the tracking substantially. Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results. Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.

@@

Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly when one flower occludes another.

Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max disappeared is set to zero. Introducing a value for this parameter can deal with the problem of false negatives.

@@

>>>>

A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track. Often simple statistics are used to describe the flowering phenology of a population of plants, e.g. onset or peak of flowering. Similarly, flower visitation rates are based on crude observation of flower populations. Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individuals. Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series. We compare the results of automatic tracking with manual tracking. Individual flower tracking could facilitate finegrained studies on plant reproductive success.

“and we show that performance achieved by following automatic tracking can be reliable enough to estimate X, Y, Z.”
---

@@

In this paper, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology. We present and evaluate an automatic flower tracking and filtering algorithm. Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. We therefore base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.

DeepSORT and other tracking frameworks leverage convolutional neural networks (CNNs) for object classification.

But in our case, flowers are not distinguishable and one flower may look less alike between two frames than two flowers in a single frame.

For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach. The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors. The approach allows for confident upscaling of the method. I.e., when applying the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.

@@

`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and assigned each individual flower a unique ID. These annotations constitute our ground truth tracks.

@@

The tracking algorithm tracks objects based on distances between centroids of bounding boxes. The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy. The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers. It is important to note, however, that the tracking algorithm could be used to track any objects. The algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame). The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked. The method is fast, however. Tracking of a series containing 85 objects ran at ~0.02 seconds per frame on a standard laptop.

Several issues complicate the task of tracking individual flowers through the time-lapse series. First, as the wind shifts, the flower heads changes direction. This can happen instantaneously (i.e., between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other. With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.

Second, when the location of the flowers changes they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view. This can cause a track to be lost and a new track erroneously being initiated when the flower reappears. However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed. The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area. Concurrently, this deals with potential false negatives. If a given flower has not been detected in a few frames, the track will not be lost. The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.

@@

# EDIT FIGURES: Zoom in, bigger dots, combine into one figure.
#
![](../figures/figure_1.png){ width=100% }

@@

#
### Input data for the tracking algorithm

The input data of

Calculating tracking accuracy requires a ground truth ID for each bounding box.

This data structure is the same that is output from manual object annotations and object detection methods with CNNs.

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100. Note that set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.

`r word_comment("I haven't included these results. Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground truth tracks. First, we calculated the largest distance between any two points within any track for any flower in each of the five series. Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.

#
## Evaluating tracking performance

The optimal way of quantifying tracking performance depends on the goal of the tracking. To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct. To derive flowering length, in theory we just need to be able to connect the first and the last image of a flower and can ignore the intermediate points. Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.

The multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms. Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal.

@@

We base our filtering on a density-based clustering of the centroid of a geometry of each track. The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid. For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line. For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle. For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon. For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.

We then apply the filtering based on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches. The filtering is done in two steps. First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.

We identify a good value for eps by...

Second, all tracks that were not assigned to a unique cluster are removed. Tracks that are spatially isolated remains. We evaluate the tracking accuracy of the remaining tracks. We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.

@@

The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig. 4. Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size and max distance, respectively).

@@

We applied our tracking algorithm on 3179 combinations of the three parameters. Using the parameters increased the performance of the tracking substantially. Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results. Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.

We have demonstrated that our tracking algorithm with the three tracking parameters and the subsequent filtering algorithm can substantially increase accuracy of the tracking of flowers despite the fact that they break the assumptions of general tracking methods.

In a naive setting, a priori knowledge about which settings to use is not available.

As with any tracking method, accuracy can only be assessed based on visual ... or more ... based on a subsample of tracks for whih a ground truth exists.

We recommend to perform a parameter test as demonstrated here on subset of ground truth tracks to derive optimal parameters for application of the method in a new setting.

@@

Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly, when one flower occludes another.

Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max gap size is set to zero. Introducing a value for this parameter can deal with the problem of false negatives.

@@

**Alekos: I think it would be good to add a short description or comment on what each of the mentioned parameters is used for (e.g. why did you need to change its value, how this affects the performance of the method on the challenges in your dataset compared to standard computer vision tracking datasets).**
